---
title: "p8105_hw2_xq2270"
author: "Xiaohan Qiang"
date: "2026-10-06"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  echo = TRUE,
  results = "hide",
  warning = FALSE,
  message = FALSE
)
```
# Problem 1
```{r}
subway_df = 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```

The cleaned subway dataset contains `r nrow(subway_df)` observations and
`r ncol(subway_df)` variables. It includes information on subway line and
station name, station latitude and longitude, routes served, entrance type,
entry and vending status, and ADA compliance. During data cleaning, I
standardized variable names, retained only the variables relevant to the
analysis, and converted the `entry` variable from character values ("YES"
and "NO") to a logical variable. The dataset is not fully tidy because the
routes served by each station are stored across multiple columns
(`route1` through `route11`) rather than in separate route number and route
name variables.

```{r}
distinct_stations = 
  subway_df |>
  distinct(line, station_name)

nrow(distinct_stations)
```
There are `r nrow(distinct_stations)` distinct subway stations in the dataset.

```{r}
ada_stations = 
  subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name)

nrow(ada_stations)
```
There are `r nrow(ada_stations)` ADA-compliant stations.

```{r}
no_vending_entry = 
  subway_df |>
  filter(vending == "NO") |>
  summarise(
    proportion_entry = mean(entry)
  )

no_vending_entry
```
Among station entrances and exits without vending, approximately
`r round(no_vending_entry$proportion_entry * 100, 1)`% allow entrance.

```{r}
subway_long = 
  subway_df |>
  mutate(
    across(route1:route11, as.character)
  ) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    values_drop_na = TRUE
  )
```
```{r}
a_train_stations = 
  subway_long |>
  filter(route_name == "A") |>
  distinct(line, station_name)

a_train_ada_stations = 
  subway_long |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name)
```
After reorganizing the route variables, there are
`r nrow(a_train_stations)` distinct stations that serve the A train.
Of these, `r nrow(a_train_ada_stations)` stations are ADA compliant.

# Problem 2

# Problem 3